---
title: "Sta 523 - Homework 5"
author: Team 08
output: rmarkdown::html_document
runtime: shiny
---

### Setup

```{r, include = FALSE, message = FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(repos="https://cran.rstudio.com" )
install.packages("googleway")
``````


```{r, message = FALSE}
# Load any necessary packages here

library(shiny)
library(jsonlite)
library(dplyr)
library(rvest)
library(stringr)
library(googleway)
```

<br/><br/>

### Task 1 - Getting data from Dark Sky

<!-- Include your write up here and or below -->

When user does not choose a date, a 48-hour forecast is provided by our app. we first convert the inputs to valid url for Forecast Request, and obtain the JSON data from the website. We then convert it to R object, and get the hourly data elements in the list, which is by default a data frame with 49 rows, 1 row for each hour.

When user does choose a past date, our app shows the weather for two days prior and following the chosen date. For each day, we calculate the right date and convert the inputs to valid url for Time Machine Request, and obtain the JSON data from the website. We then convert it to R object, and get the hourly data elements in the list, which is by default a data frame with 24 rows, 1 row for each hour. The four data frames are then combined into one full data frame with 96 rows. Cells without data are filled with NA's as a result of row bind.

```{r}
get_darksky = function(key, lat, long, date = NULL){
  base_url = "https://api.darksky.net/forecast"
  day_sec = 60*60*24
  
  # Get data from web
  if(is.null(date)){               ##Forecast Request
    hourly = fromJSON(paste(base_url,key,paste(lat,long,sep=","),sep="/"))$hourly$data
  }else{                           ##Time Machine Request
    hourly = lapply(-2:1, function(i)
      fromJSON(paste(base_url,key,
                     paste(lat,long,date+i*day_sec,sep=","),sep="/"))$hourly$data)%>% 
      bind_rows()
  }
  
  return(hourly)
}
```



<br/><br/>

### Task 2 - Prediction Locations

<!-- Include your write up here and or below -->

We first scrap US city location information from the table on wikipedia by getting the first sortable wikitable on that page, store it as raw table, and rename the columns accordingly.

We delete the rank and imperial unit columns from the raw table, and convert each variables to the right form. We clean the `City` column by removing any footnote references. The `State` column is already clean. The commas in the `census_2010` column is dropped and the values converted to numeric type. The data frame is then filtered by the constrain that the city needs to have more than 500,000 residents according the the census in 2010. The noise in the `change` column is dropped and we convert it from a percentage form string to a decimal number of numeric type. The units and commas in the `land_area_2016` and `pop_density_2016` are dropped and we convert them to numeric type. Note that for `land_area_2016` and `pop_density_2016` we are using the metric units. The `latitude` and `longitude` are extracted from `Location` and converted to numeric type. The `Location` column is then deleted.

```{r}
# Create your location data frame here
us_locs = data.frame()

# Get raw table from web
page = read_html("https://en.wikipedia.org/wiki/List_of_United_States_cities_by_population")
raw_table = page %>% html_nodes('table[class="wikitable sortable"]') %>% 
{.[1]} %>% html_table() %>% {.[[1]]} 
names(raw_table) = c("rank", "City", "State", "estimate_2016", "census_2010", "change",
                     "land_area_2016_sq_mi", "land_area_2016", "pop_density_2016_sq_mi",
                     "pop_density_2016", "Location")

# Use metric units
us_locs = raw_table %>% select(-rank, -land_area_2016_sq_mi, -pop_density_2016_sq_mi) %>% 
  mutate(City = str_replace(City, "\\[\\d+\\]",""),
         estimate_2016 = str_replace_all(estimate_2016,",","") %>% as.numeric(),
         census_2010 = str_replace_all(census_2010,",","") %>% as.numeric()) %>% 
  filter(census_2010>500000 ) %>% 
  mutate(change = str_replace_all(change, "\\d+???|%","") %>% 
           str_replace("???","-") %>% as.numeric()/100,
         land_area_2016 = str_replace_all(land_area_2016, ",|\\s[\\s,\\S]*","") %>% 
           as.numeric(),
         pop_density_2016 = str_replace_all(pop_density_2016, ",|\\/[\\s,\\S]*","") %>% 
           as.numeric(),
         latitude = str_replace_all(Location, 
                                    "[\\s,\\S]*?/[\\s,\\S]*?/\\s|;[\\s,\\S]*$","") %>% 
           as.numeric(),
         longitude = str_replace_all(Location,"[\\s,\\S]*?;\\s","") %>% 
           str_extract("-\\d*.\\d*") %>% as.numeric(),
         Location = NULL)
```
 
 <br/><br/>

### Task 3 - Shiny Predictions


<!-- Include your write up here and or below -->
The purpose of our app is to plot the temperature and a 2nd weather-related quantity of chosen city and date. The UI allow users to choose city from a list, which is obtained in task 2, a day from 1970-1-1 to current date, and a 2nd quantity. The list of 2nd quantity consists of all the numeric variables of the JSON data of selected city and date. When a city is selected, the server will extract its latitude and longitude from the table of US cities. The Server then use the latitude, longitude and date to send a request to darksky API by `get_darksky()` function. Darksky API will then sent back the weather data which we later use for the plot. The `reactive()`stores variables `lat`, `long`, `timezone`, `date`, `hourly`, `quants`, the 2nd quantity list `names.display`, and title as reactive functions, which make them easier to update later in render functions. The `Observeevent()` in the Server will observe the input of city and date, and update the list of 2nd quantity in UI based on available variables in the JSON file. If the chosen 2nd quantity is not available for current city and date, the validating code will output a error message.

There is a `plotOutput` and a `textOutput` in the `mainPanel`. The plot shows how temperature will change from the current time to 48 hours later if a forecast request is made(no date input), and the observed change of temperature if a time machine request is made. If user has chosen one in the `sidePanel`, the plot will also display a 2nd quantity, such as pressure or humidity. We scale the plot of 2nd quantity in order to make the range of the whole plot reasonable. The unscaled axis of the 2nd quantity is shown on the right the plot, so users can read the original numbers from the plot easily.

```{r echo=FALSE}
# Modify this default shiny app

## Initiate Default using first city in us_locs, Forcast request
## In order to get choice of all other posible quantities for panel
default.city = us_locs$City[1]
hourly = get_darksky("dfb209f3d5355ec88b0a429d80eb9a2a", 
                     us_locs$latitude[us_locs$City==default.city], 
                     us_locs$longitude[us_locs$City==default.city], NULL)
quants = hourly[sapply(hourly,is.numeric)] %>%
  select(-time, -temperature) %>% names()
names.display = gsub("([a-z])([A-Z])", "\\1 \\2",quants) %>% str_to_title()

## A data frame for matching all possible units
unit.df = data_frame(temperature = "Fahrenheit",
           precipIntensity = "Millimeters per hour",
           precipProbability = "percentage", 
           apparentTemperature = "Fahrenheit",
           dewPoint = "Fahrenheit", humidity = "percentage",
           pressure = "millibars", windSpeed = "miles per hour",
           windBearing = "degrees", cloudCover = "percentage",
           uvIndex = "index level", visibility = "mile",
           precipAccumulation = "inches", ozone = "DU",
           windGust = "miles per hour")




shinyApp(
  ########## ui
  ui = fluidPage(
     ## Title
     titlePanel("Local Weather Hourly Forecast"),

     ## Layout
     sidebarLayout(
        ######## Sidebar Panel
        sidebarPanel(
          ## Location input
          h4("Location:"),
          selectInput(inputId='city', labe='Choose City', choices = us_locs$City, 
                      selected=us_locs$City[1]), #Default city New York
          br(),
          
          ## Second quantity input
          h4("Second quantity:"),
          checkboxInput("askSecond", "Want a second quantity?"),
          conditionalPanel(
            "input.askSecond == true",
            selectInput("second","Choose second quantity", 
                        choices = names.display, 
                        selected = names.display[1])),
          ######Note: Need to update names.display
          br(),
          
          ## Date input
          h4("Date:"),
          checkboxInput("askDate", "Choose Date for Time Machine Request?"),
          conditionalPanel(
            "input.askDate == true",
            dateInput("date", "Choose Date", value = Sys.Date(), 
                      min = "1970-1-1", max = Sys.Date())
          )
        ),
        
        ######### Main Panel
        mainPanel(
          plotOutput('plot'),
          htmlOutput('text')
        )
     )
  ),
  
  
  ######### server
  server = function(input, output, session) {
    ###### Reactives
    lat = reactive({us_locs$latitude[us_locs$City==input$city]})
    long = reactive({us_locs$longitude[us_locs$City==input$city]})
    timezone = reactive({google_timezone(c(lat(), long()),key = NULL)$timeZoneId})
    
    date = reactive({
      if(input$askDate){
        output = as.numeric(as.POSIXct(as.character(input$date), tz = timezone()))
      }else{output = NULL}
      output
    })
    
    title.name = reactive({
      if(input$askDate){
        output = "Hourly Forecast for the Two Days Prior and Two Days Following in "
      }else{
        output =  "Hourly Forecast for the Next Two Days in "}
      output
    })
    
    hourly = reactive({
      get_darksky(key = "dfb209f3d5355ec88b0a429d80eb9a2a",lat(),long(),date())
    })
      
    quants = reactive({
      hourly = hourly()
      hourly[sapply(hourly,is.numeric)] %>% select(-time, -temperature) %>% names()
    })
      
    names.display = reactive({
      gsub("([a-z])([A-Z])", "\\1 \\2",quants()) %>% str_to_title()
    })
      

    
    ##### Update 2nd variable panel
    observeEvent({
      input$city
      input$askDate
      input$date
    },{
      names.display = names.display()
      
      # If selected 2nd quantity not valid, 
      # change selected to 1st element of new panel
      if(input$second %in% names.display){
        updateSelectInput(session, "second", 
                        choices = names.display,
                        selected = input$second)
      }else{
        updateSelectInput(session, "second", 
                        choices = names.display,
                        selected = names.display[1])
      }
    })
    
    
    ##### output for plot
    output$plot = renderPlot({
      timezone = timezone()
      title.name = title.name()
      hourly = hourly()
      
      ## First check if 2nd quantity is valid
      ## return if not
      quants = quants()
      names.display = names.display()
      if(!input$second %in% names.display)
        return()

      # Plot temperature
      plot = ggplot(hourly, aes(x = as.POSIXct(time, origin="1970-01-01 00:00:00",
                                        tz = timezone, format = "%Y-%m-%d")))+
        geom_line(aes(y = temperature, color='Temperature'))+
        labs(title=paste0(title.name, input$city),
             x ="Date time", y = paste0("Temperature/",unit.df[["temperature"]]))
      colors = c("#9999CC")
      # Add another y to plot if needed
      if(input$askSecond){
        quant2 = quants[names.display == input$second]
        scale = max(hourly[[quant2]])/max(hourly$temperature)
        plot = plot + 
          geom_line(aes(y = hourly[[quant2]]/scale, colour = input$second)) +
          scale_y_continuous(sec.axis = sec_axis(~.*scale, 
                        name = paste0(input$second,"/",unit.df[[quant2]])))
        colors = c("#CC6677", "#9999CC")
      }
      plot + scale_color_manual(values=colors)
    })
    
    ###### output for text
    output$text <- renderUI({
      HTML(paste(paste0(input$city, ":"), paste("Latitude:",lat()),
                 paste("Longitude:",long()), sep = '<br/>'))
    })
    
  }
)
```